\singlespacing \vspace{-1.6cm} \LARGE Trade-offs among plant reproductive traits determine interactions with floral visitors

### 2022-04-22 Manuscript/Lanuza_et_al_2021.tex
edits

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{float}
\floatplacement{figure}{H}
\usepackage[skip=3pt]{caption}
\captionsetup[figure]{labelformat=empty}
\usepackage{setspace}
\usepackage{titlesec}
\titlespacing{\title}{0pt}{\parskip}{-\parskip}
\usepackage[style=nature]{biblatex}
\usepackage[labelformat = empty]{caption}
\usepackage{pdflscape}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

@@

Despite the astonishing diversity of floral structures among flowering plants (\protect\hyperlink{ref-barrett2002}{Barrett 2002}; \protect\hyperlink{ref-schiestl2013}{Schiestl \& Johnson 2013}) and their importance in shaping plant-pollinator interactions (\protect\hyperlink{ref-fenster2004}{Fenster \emph{et al.} 2004}; \protect\hyperlink{ref-dellinger2020}{Dellinger 2020}), a unified framework that describes the major plant reproductive trade-offs is currently lacking (\protect\hyperlink{ref-roddy2021}{Roddy \emph{et al.} 2021}). In addition, macroecological studies that investigate plant reproductive traits are scarce (\protect\hyperlink{ref-baude2016}{Baude \emph{et al.} 2016}; \protect\hyperlink{ref-munoz2016}{Munoz \emph{et al.} 2016}; \protect\hyperlink{ref-grossenbacher2017}{Grossenbacher \emph{et al.} 2017}; \protect\hyperlink{ref-moeller2017}{Moeller \emph{et al.} 2017}) and consequently, there is poor understanding of how reproductive traits drive interactions with pollinators at large scales (\protect\hyperlink{ref-sargent2008}{Sargent \& Ackerly 2008}; \protect\hyperlink{ref-rech2016}{Rech \emph{et al.} 2016}; \protect\hyperlink{ref-salguero2016}{Salguero-Gómez \emph{et al.} 2016}; \protect\hyperlink{ref-ruger2018}{Rüger \emph{et al.} 2018}). Linking the plant's position in trait-space with the different pollinator groups could help to improve our understanding of plant-pollinator associations (\protect\hyperlink{ref-dehling2016}{Dehling \emph{et al.} 2016}). Further, there is increasing interest in understanding drivers of plant-pollinator interactions using trait-based approaches (\protect\hyperlink{ref-fenster2004}{Fenster \emph{et al.} 2004}; \protect\hyperlink{ref-rosas2014}{Rosas-Guerrero \emph{et al.} 2014}) and trait-matching analyses (\protect\hyperlink{ref-stang2009}{Stang \emph{et al.} 2009}; \protect\hyperlink{ref-bartomeus2016}{Bartomeus \emph{et al.} 2016}). However, plant reproductive traits have been overlooked beyond highly specialised pollination systems (\protect\hyperlink{ref-dellinger2020}{Dellinger 2020}) and remains unclear how specific plant reproductive biology traits (e.g., mating or compatibility system) influence plant-pollinator associations (\protect\hyperlink{ref-tur2013}{Tur \emph{et al.} 2013}; \protect\hyperlink{ref-devaux2014}{Devaux \emph{et al.} 2014}).

Species can optimise their fitness through various life-history traits, yet trade-offs among those traits constrain the range of potential strategies that a species can use. With the recent availability of large trait databases (e.g., TRY \protect\hyperlink{ref-kattge2011}{Kattge \emph{et al.} 2011}; and COMPADRE \protect\hyperlink{ref-salguero2015}{Salguero-Gómez \emph{et al.} 2015}), plant ecological strategies are being increasingly examined, and are facilitating the identification of global patterns and constraints in plant form and function (\protect\hyperlink{ref-diaz2016}{Díaz \emph{et al.} 2016}; \protect\hyperlink{ref-salguero2016}{Salguero-Gómez \emph{et al.} 2016}; \protect\hyperlink{ref-bruelheide2018}{Bruelheide \emph{et al.} 2018}; \protect\hyperlink{ref-carmona2021}{Carmona \emph{et al.} 2021}). However, most studies have focused on vegetative traits such as leaf (\protect\hyperlink{ref-wright2004}{Wright \emph{et al.} 2004}), wood (\protect\hyperlink{ref-chave2009}{Chave \emph{et al.} 2009}), or root (\protect\hyperlink{ref-laughlin2021}{Laughlin \emph{et al.} 2021}) trade-offs with little or no attention given to reproductive traits (\protect\hyperlink{ref-evojtko2020}{E-Vojtkó \emph{et al.} 2020}; \protect\hyperlink{ref-roddy2021}{Roddy \emph{et al.} 2021}) which are critical to plant life strategies that shape interactions with pollinators and ultimately determine plant reproductive success. For instance, short lived versus perennial species tend to have low versus high levels of outcrossing, respectively, (\protect\hyperlink{ref-barrett2003}{Barrett 2003}; \protect\hyperlink{ref-moeller2017}{Moeller \emph{et al.} 2017}) and outcrossing levels are positively correlated with flower size (\protect\hyperlink{ref-goodwillie2010}{Goodwillie \emph{et al.} 2010}). In addition, the presence of costly rewards (e.g., pollen or nectar) and showy flowers or floral displays can only be understood through consideration of plant species' reliance upon animal pollination (pollinator dependence) and its role in attracting pollinators (\protect\hyperlink{ref-ollerton2011}{Ollerton \emph{et al.} 2011}; \protect\hyperlink{ref-rodger2021}{Rodger \emph{et al.} 2021}). However, it is still unknown to what extent these different reproductive compromises determine plant-pollinator interactions.

Several studies have identified links between plant traits and plant-pollinator network properties (\protect\hyperlink{ref-bartomeus2013}{Bartomeus 2013}; \protect\hyperlink{ref-olito2015}{Olito \& Fox 2015}; \protect\hyperlink{ref-rowe2020}{Rowe \emph{et al.} 2020}). Moreover, plant traits can define species' network roles (e.g., specialists vs generalists; \protect\hyperlink{ref-lazaro2013}{Lázaro \emph{et al.} 2013}; \protect\hyperlink{ref-tur2013}{Tur \emph{et al.} 2013}). For example, plant species that occupy reproductive trait space extremes are more likely to exhibit higher levels of specialisation and be more reliant on the trait-matching with pollinators (\protect\hyperlink{ref-junker2013}{Junker \emph{et al.} 2013}; \protect\hyperlink{ref-coux2016}{Coux \emph{et al.} 2016}). Morphological matching between plant and floral visitors often determines plant-pollinator interactions, and can thus strongly influence interaction network structure (\protect\hyperlink{ref-stang2009}{Stang \emph{et al.} 2009}; \protect\hyperlink{ref-ibanez2012}{Ibanez 2012}). Importantly, the combination of traits has shown to increase the predictive power of the network interactions (\protect\hyperlink{ref-eklof2013}{Eklöf \emph{et al.} 2013}) and by considering the position in the multidimensional trait space we can determine species' roles within plant-pollinator interaction networks (\protect\hyperlink{ref-dehling2016}{Dehling \emph{et al.} 2016}). Nonetheless, we know little if those patterns generally studied at the community level are representative of wider macroecological scales.

Here, we aim to explore the potential trade-offs among reproductive traits and how these influence plant-pollinator interactions. First, we identify the major axes of reproductive trait variation and trade-offs that determine plant form and function. Second, we investigate how plant species' position in trait-space influence interactions with floral visitors. Finally, we investigate how both the main axes of trait variation, and individual traits, influence plant species' roles within networks using a set of complementary interaction network metrics (i.e., visitation rate, normalized degree and specialization).

@@

We selected 28 studies from 18 different countries that constituted a total of 64 plant-pollinator networks. These studies recorded plant-pollinator interactions in natural systems and were selected so that we had broad geographical representation. Although these studies differ in sampling effort and methodology, all studies provided information about plant-pollinator interactions (weighted and non-weighted), which we used to build a database of plant species that are likely to be animal pollinated. Many of these networks are freely available either as published studies (\protect\hyperlink{ref-olesen2007}{Olesen \emph{et al.} 2007}; \protect\hyperlink{ref-fortuna2010}{Fortuna \emph{et al.} 2010}; \protect\hyperlink{ref-carvalheiro2014}{Carvalheiro \emph{et al.} 2014}) or available in online archives (e.g., The Web of Life, \protect\hyperlink{ref-fortuna2010}{Fortuna \emph{et al.} 2010}; and Mangal, \protect\hyperlink{ref-poisot2016}{Poisot \emph{et al.} 2016}). In total, our network dataset (Table S1) constituted 60 weighted (visitation rate) and 4 unweighted (presence-absence of the interaction) networks, each sampled at a unique location and year, as well as eight meta-webs where interactions were pooled across several locations and multiple years.

@@

All species names, genera, families and orders were retrieved and standardized from the taxonomy data sources NCBI (\url{https://www.ncbi.nlm.nih.gov/taxonomy}) for plants and ITIS (\url{https://www.itis.gov/}) for pollinators, using the R package \emph{taxize} (\protect\hyperlink{ref-chamberlain2020}{Chamberlain \emph{et al.} 2020}). We filled the `not found' searches manually using \url{http://www.theplantlist.org/} and \url{http://www.mobot.org/} for plants and \url{http://www.catalogueoflife.org/} for floral visitors.

@@

We selected a total of 19 different functional traits that included both reproductive and vegetative traits (see Table 1 and Supplementary Information). From these, 16 were reproductive traits (13 floral and 3 reproductive biology traits) and were selected based on their relevance to plant reproduction and data availability. The 3 remaining traits, were vegetative traits that are commonly used to characterize the global spectrum of plant form and function and represent the fast-slow continuum of trait variation (e.g., short-lived versus perennial species). For each plant species, we undertook an extensive literature and online search across a wide range of resources (plant databases, online floras, books, journals and images). From a total of 30,120 cells (20 columns × 1,506 species) we were able to fill 24,341 cells (80.8\% of the dataset, see Fig. S1 for missing values information for each trait).

@@

We calculated the phylogenetic distance between different plant species using the function \emph{get\_tree} from the package \emph{rtrees} (\url{https://github.com/daijiang/rtrees}), which downloads phylogenetic distances from the extended R implementation of the Open Tree of Life (\protect\hyperlink{ref-smith2018}{Smith \& Brown 2018}; \protect\hyperlink{ref-jin2019}{Jin \& Qian 2019}).

@@

Trait missing values were imputed with the function \emph{missForest} (\protect\hyperlink{ref-stekhoven2012}{Stekhoven \& Bühlmann 2012}) which allows imputation of data sets with continuous and categorical variables. We accounted for the phylogenetic distance among species on the imputation process by including the eigenvectors of a principal component analysis of the phylogenetic distance (PCoA) which has been shown to improve the performance of \emph{missForest} (\protect\hyperlink{ref-penone2014}{Penone \emph{et al.} 2014}). To extract the eigenvectors, we used the function \emph{PVRdecomp} from the package \emph{PVR} (\protect\hyperlink{ref-santos2018}{Chamberlain \emph{et al.} 2018}) based on a previous conceptual framework that considers phylogenetic eigenvectors (\protect\hyperlink{ref-diniz-filho2012}{Diniz-Filho \emph{et al.} 2012}). We conducted two different imputations, one for the full set of species (1,506 species, 5.79\% of missing values) excluding nectar and pollen traits because of the high percentage of missing values (Fig. S1) and a second one for the subset of species with data for pollen per flower and microliters of nectar (755 species, 8.01\% of missing values).

@@

We explored the trade-offs between the different quantitative plant traits with a phylogenetically informed Principal Component Analysis (pPCA). We did not include the quantitative variables of flower length and inflorescence width because they were highly and moderately correlated to flower width respectively (Pearson's correlation = 0.72, \emph{P} \textless{} 0.01 and Pearson's correlation = 0.36, \emph{P} \textless{} 0.01), and thus we avoided overemphasizing flower size on the spectrum of trait variation. Although qualitative traits were not included in the dimensionality reduction analysis, we also investigated the association of the different qualitative traits with the main axes of trait variation. Prior to the analyses, we excluded outliers and standardized the data. Due to the high sensitivity of dimensionality reduction to outliers, we excluded values within the 2.5th--97.5th percentile range (\protect\hyperlink{ref-legendre2012}{Legendre \& Legendre 2012}), and thus our final dataset had 1,236 species. Then, we log transformed the variables to reduce the influence of outliers and z-transformed (X= 0, SD=1) so that all variables were within the same numerical range. We performed the pPCA using the function \emph{phyl.pca} from the package \emph{phytools} (\protect\hyperlink{ref-revell2012}{Revell 2012}) with the method lambda (\(\lambda\)) that calculates the phylogenetic correlation between 0 (phylogenetic independence) and 1 (shared evolutionary history) and we implemented the mode covariance because values for each variables were on the same scale following transformation (\protect\hyperlink{ref-abdi2010}{Abdi \& Williams 2010}). Moreover, to corroborate that our imputation of missing values did not affect our results, we conducted a pPCA on the full dataset without missing values (Fig. S2). We found little difference between the explained variance with the imputed dataset (51.08\%) and the dataset without missing values (52.87\%). In addition, the loadings on each principal component had a similar contribution and correlation patterns, with the exception of plant height which showed slight variations between the imputed and non-imputed dataset. Finally, we conducted an additional phylogenetic informed principal component analysis for the subset of species with pollen and nectar quantity. For this, we included all quantitative traits considered in the main pPCA plus pollen grains and microlitres of nectar per flower.

@@

We calculated the phylogenetic signal of the different quantitative traits on the imputed dataset with the full set of species (N = 1,506) with the package \emph{phytools} (\protect\hyperlink{ref-revell2012}{Revell 2012}) and we used Pagel's \(\lambda\) as a measurement of the phylogenetic signal. However, for pollen and nectar traits, phylogenetic signal was calculated only on the subset of species that had quantitative information for these traits (N = 755).

@@

Analyses were conducted on the subset of 60 weighted networks sampled in a unique flowering season and site, which included 556 plant and 1,126 pollinator species. These networks were analysed in their qualitative and quantitative form. First, we analysed the binary version of these weighted networks with presence-absence information that assumes equal weight across interactions. Second, we analysed the untransformed weighted networks with visitation rate that accounts for the intensity of the interaction. Although floral visitors are not always pollinators and visitation rate does not consider each pollinator species efficiency (\protect\hyperlink{ref-ballantyne2015}{Ballantyne \emph{et al.} 2015}), visitation rate can provide valuable information of the contribution of floral visitors to pollination (\protect\hyperlink{ref-vazquez2005}{Vázquez \emph{et al.} 2005}, \protect\hyperlink{ref-vazquez2012}{2012}). In total, our network dataset (excluding meta-webs and non-weighted networks) included 2,256 interactions of bees with plants, 1,768 non-syrphid-Diptera interactions, 845 syrphids interactions, 437 Lepidoptera interactions, 432 Coleoptera interactions and 362 non-bee-Hymenoptera interactions. Sampling methods varied across networks but this was accounted for in analyses by considering them in the random effects of the modelling process. All analyses were conducted in R version \emph{4.0.3}.

@@

We used Bayesian modelling (see below for details) to explore the effect of floral visitor groups and the main axes of trait variation (pPCA with imputed dataset) on both qualitative and quantitative floral interactions per plant species. For this, we divided floral visitors into six main guilds that differ in life form, behaviour and are likely to play a similar ecological role: (i) bees (Hymenoptera-Anthophila), (ii) non-bee-Hymenoptera (Hymenoptera-non-Anthophila), (iii) syrphids (Diptera-Syrphidae), (iv) non-syrphid-Diptera (Diptera-non-Syrphidae), (v) Lepidoptera and (vi) Coleoptera. Moreover, because the guild of bees was the most represented group with 2,256 records and had the highest frequency of visits, we also explored presence-absence of the interaction and visitation rate of the main bee families (Andrenidae, Apidae, Colletidae, Halictidae and Megachilidae) on the trait space. In addition, we found that \emph{Apis mellifera} was the floral visitor with the largest proportion of records counted (7.55\% of the total). This finding is consistent with previous research showing that \emph{A. mellifera} was the most frequent floral visitor in a similar dataset of 80 plant-pollinator networks in natural ecosystems (\protect\hyperlink{ref-hung2018}{Hung \emph{et al.} 2018}). Hence, to control for the effect of \emph{A. mellifera} on the observed visitation patterns of bees, we conducted an analogous analysis with presence-absence of the interaction and visitation rate excluding \emph{A. mellifera}. We found that \emph{A. mellifera}, was partly driving some of the observed trends on PC1 (Fig. S3). However, we did not detect major differences on PC2 and PC3.

>>>>

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{float}
\floatplacement{figure}{H}
\usepackage[skip=3pt]{caption}
\captionsetup[figure]{labelformat=empty}
\usepackage{setspace}
\usepackage{titlesec}
\titlespacing{\title}{0pt}{\parskip}{-\parskip}
\usepackage[style=nature]{biblatex}
\usepackage[labelformat = empty]{caption}
\usepackage{pdflscape}
\usepackage[T1]{fontenc}
\DeclareUnicodeCharacter{0301}{*************************************}
\DeclareUnicodeCharacter{1EF3}{*************************************}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

@@

Flowering plants have an astonishing diversity of floral structures (\protect\hyperlink{ref-barrett2002}{Barrett, 2002}; \protect\hyperlink{ref-schiestl2013}{Schiestl \& Johnson, 2013}) that shape plant-pollinator associations (\protect\hyperlink{ref-dellinger2020}{Dellinger, 2020}; \protect\hyperlink{ref-fenster2004}{Fenster et al., 2004}). However, not all reproductive trait combinations are possible due to evolutionary and ecological constraints (\protect\hyperlink{ref-salguero2016}{Salguero-Gómez et al., 2016}). Despite the recent advances in the theoretical understanding of the major floral trade-offs {[}e.g., floral economic spectrum; \protect\hyperlink{ref-roddy2021}{Roddy et al.} (\protect\hyperlink{ref-roddy2021}{2021}){]}, we lack empirical evidence that helps to elucidate the spectrum of reproductive trait variation at a global scale as done recently for other vegetative and physiological traits (\protect\hyperlink{ref-diaz2016}{Díaz et al., 2016}; \protect\hyperlink{ref-onoda2017}{Onoda et al., 2017}).

Although there is an increasing number of macroecological studies that investigate plant reproductive traits (\protect\hyperlink{ref-baude2016}{Baude et al., 2016}; \protect\hyperlink{ref-grossenbacher2017}{Grossenbacher et al., 2017}; \protect\hyperlink{ref-moeller2017}{Moeller et al., 2017}; \protect\hyperlink{ref-munoz2016}{Munoz et al., 2016}), we still have poor understanding of how reproductive traits drive interactions with pollinators at large scales (\protect\hyperlink{ref-rech2016}{Rech et al., 2016}; \protect\hyperlink{ref-ruger2018}{Rüger et al., 2018}; \protect\hyperlink{ref-salguero2016}{Salguero-Gómez et al., 2016}; \protect\hyperlink{ref-sargent2008}{Sargent \& Ackerly, 2008}). Because the species' morphology can determine the species' functional role, by linking the plant's position in the reproductive trait-space with the different pollinator groups (as done similarly in \protect\hyperlink{ref-dehling2016}{Dehling et al., 2016}), we could help to improve our understanding of plant-pollinator associations. Further, there is increasing interest in understanding drivers of plant-pollinator interactions using trait-based approaches (\protect\hyperlink{ref-fenster2004}{Fenster et al., 2004}; \protect\hyperlink{ref-rosas2014}{Rosas-Guerrero et al., 2014}) and trait-matching analyses (\protect\hyperlink{ref-bartomeus2016}{Bartomeus et al., 2016}; \protect\hyperlink{ref-stang2009}{Stang et al., 2009}). However, plant reproductive traits have been overlooked for most pollination systems (\protect\hyperlink{ref-dellinger2020}{Dellinger, 2020}) and remains unclear how specific plant reproductive biology traits (i.e., mating or compatibility system) influence plant-pollinator associations (\protect\hyperlink{ref-devaux2014}{Devaux et al., 2014}; \protect\hyperlink{ref-tur2013}{Tur et al., 2013}).

Species can optimise their fitness through various life-history traits, yet trade-offs among those traits constrain the range of potential strategies that a species can use. With the recent availability of large trait databases (e.g., TRY \protect\hyperlink{ref-kattge2011}{Kattge et al., 2011}; and COMPADRE \protect\hyperlink{ref-salguero2015}{Salguero-Gómez et al., 2015}), plant ecological strategies are being increasingly examined, and are facilitating the identification of global patterns and constraints in plant form and function (\protect\hyperlink{ref-bruelheide2018}{Bruelheide et al., 2018}; \protect\hyperlink{ref-carmona2021}{Carmona et al., 2021}; \protect\hyperlink{ref-diaz2016}{Díaz et al., 2016}; \protect\hyperlink{ref-salguero2016}{Salguero-Gómez et al., 2016}). However, most studies have focused on vegetative traits such as leaf (\protect\hyperlink{ref-wright2004}{Wright et al., 2004}), wood (\protect\hyperlink{ref-chave2009}{Chave et al., 2009}), or root (\protect\hyperlink{ref-laughlin2021}{Laughlin et al., 2021}) trade-offs with little or no attention given to reproductive traits (\protect\hyperlink{ref-evojtko2020}{E-Vojtkó et al., 2020}; \protect\hyperlink{ref-roddy2021}{Roddy et al., 2021}) which are critical to plant life strategies that shape interactions with pollinators and ultimately determine plant reproductive success. For instance, short lived versus perennial species tend to have low versus high levels of outcrossing, respectively, (\protect\hyperlink{ref-barrett2003}{Barrett, 2003}; \protect\hyperlink{ref-moeller2017}{Moeller et al., 2017}) and outcrossing levels are positively correlated with flower size (\protect\hyperlink{ref-goodwillie2010}{Goodwillie et al., 2010}). In addition, the presence of costly rewards (e.g., pollen or nectar) and showy flowers or floral displays can only be understood through consideration of plant species' reliance upon animal pollination (pollinator dependence) and its role in attracting pollinators (\protect\hyperlink{ref-ollerton2011}{Ollerton et al., 2011}; \protect\hyperlink{ref-rodger2021}{Rodger et al., 2021}). However, it is still unknown to what extent these different reproductive strategies determine plant-pollinator interactions.

Several studies have identified links between plant traits and plant-pollinator network properties (\protect\hyperlink{ref-bartomeus2013}{Bartomeus, 2013}; \protect\hyperlink{ref-olito2015}{Olito \& Fox, 2015}; \protect\hyperlink{ref-rowe2020}{Rowe et al., 2020}). Moreover, plant traits can define species' network roles (e.g., specialists vs generalists; \protect\hyperlink{ref-lazaro2013}{Lázaro et al., 2013}; \protect\hyperlink{ref-tur2013}{Tur et al., 2013}). For example, plant species that occupy reproductive trait space extremes are more likely to exhibit higher levels of specialisation and be more reliant on the trait-matching with pollinators (\protect\hyperlink{ref-coux2016}{Coux et al., 2016}; \protect\hyperlink{ref-junker2013}{Junker et al., 2013}). Morphological matching between plant and floral visitors often determines plant-pollinator interactions, and can thus strongly influence interaction network structure (\protect\hyperlink{ref-ibanez2012}{Ibanez, 2012}; \protect\hyperlink{ref-stang2009}{Stang et al., 2009}). Importantly, the combination of traits has shown to increase the predictive power of the network interactions (\protect\hyperlink{ref-eklof2013}{Eklöf et al., 2013}) and by considering the position in the multidimensional trait space we can determine species' roles within plant-pollinator interaction networks (\protect\hyperlink{ref-dehling2016}{Dehling et al., 2016}). Nonetheless, we know little if those patterns generally studied at the community level are representative of wider macroecological scales.

Here, we aim to explore the potential trade-offs among plant reproductive traits and their association with the different floral visitor guilds. First, we identify the major axes of reproductive trait variation and trade-offs that determine plant form and function. Second, we investigate the association between the plant species' position in trait-space and the different floral visitor guilds with the help of qualitative (presence-absence of interaction) and quantitative (number of visits) information. Finally, we investigate how both the main axes of trait variation, and individual traits, influence plant species' roles within networks using a set of complementary interaction network metrics (i.e., number of visits, normalized degree and specialization).

@@

We selected 28 studies from 18 different countries that constituted a total of 64 plant-pollinator networks (see Table S1 and Fig. S1). These studies recorded plant-pollinator interactions in natural systems and were selected so that we had broad geographical representation. Although these studies differ in sampling effort and methodology, all studies provided information about plant-pollinator interactions (weighted and non-weighted), which we used to build a database of plant species that are likely to be animal pollinated. Many of these networks are freely available either as published studies (e.g.,~ \protect\hyperlink{ref-carvalheiro2014}{Carvalheiro et al., 2014}; \protect\hyperlink{ref-fortuna2010}{Fortuna et al., 2010}; \protect\hyperlink{ref-olesen2007}{Olesen et al., 2007}) or available in online archives (e.g., The Web of Life, \protect\hyperlink{ref-fortuna2010}{Fortuna et al., 2010}; and Mangal, \protect\hyperlink{ref-poisot2016}{Poisot et al., 2016}). In total, our network dataset constituted 60 weighted (number of visits) and 4 unweighted (presence-absence of the interaction) networks, each sampled at a unique location and year, as well as eight meta-webs where interactions were pooled across several locations and multiple years.

@@

All species names, genera, families and orders were retrieved and standardized from the taxonomy data sources NCBI (\url{https://www.ncbi.nlm.nih.gov/taxonomy}) for plants and ITIS (\url{https://www.itis.gov/}) for pollinators, using the R package \emph{taxize} (\protect\hyperlink{ref-chamberlain2020}{Chamberlain et al., 2020}). We filled the `not found' searches manually using \url{http://www.theplantlist.org/} and \url{http://www.mobot.org/} for plants and \url{http://www.catalogueoflife.org/} for floral visitors.

@@

We selected a total of 19 different functional traits that included both reproductive and vegetative traits (see Table 1 and Supplementary Information). From these, 16 were reproductive traits (13 floral and 3 reproductive biology traits) and were selected based on their relevance to plant reproduction and data availability. The 3 remaining traits, were vegetative traits that are commonly used to characterize the global spectrum of plant form and function and represent the fast-slow continuum of trait variation (e.g., short-lived versus perennial species). For each plant species, we undertook an extensive literature and online search across a wide range of resources (plant databases, online floras, books, journals and images). From a total of 30,120 cells (20 columns × 1,506 species) we were able to fill 24,341 cells (80.8\% of the dataset, see Fig. S2 for missing values information for each trait).

@@

We calculated the phylogenetic distance between different plant species using the function \emph{get\_tree} from the package \emph{rtrees} (\url{https://github.com/daijiang/rtrees}), which downloads phylogenetic distances from the extended R implementation of the Open Tree of Life (\protect\hyperlink{ref-jin2019}{Jin \& Qian, 2019}; \protect\hyperlink{ref-smith2018}{Smith \& Brown, 2018}).

@@

Trait missing values were imputed with the function \emph{missForest} (\protect\hyperlink{ref-stekhoven2012}{Stekhoven \& Bühlmann, 2012}) which allows imputation of data sets with continuous and categorical variables. We accounted for the phylogenetic distance among species on the imputation process by including the eigenvectors of a principal component analysis of the phylogenetic distance (PCoA) which has been shown to improve the performance of \emph{missForest} (\protect\hyperlink{ref-penone2014}{Penone et al., 2014}). To extract the eigenvectors, we used the function \emph{PVRdecomp} from the package \emph{PVR} (\protect\hyperlink{ref-santos2018}{Chamberlain et al., 2018}) based on a previous conceptual framework that considers phylogenetic eigenvectors (\protect\hyperlink{ref-diniz-filho2012}{Diniz-Filho et al., 2012}). We conducted two different imputations, one for the full set of species (1,506 species, 5.79\% of missing values) excluding nectar and pollen traits because of the high percentage of missing values (Fig. S2) and a second one for the subset of species with data for pollen per flower and microliters of nectar (755 species, 8.01\% of missing values).

@@

We explored the trade-offs between the different quantitative plant traits with a phylogenetically informed Principal Component Analysis (pPCA). We did not include the quantitative variables of flower length and inflorescence width because they were highly and moderately correlated to flower width respectively (Pearson's correlation = 0.72, \emph{P} \textless{} 0.01 and Pearson's correlation = 0.36, \emph{P} \textless{} 0.01), and thus we avoided overemphasizing flower size on the spectrum of trait variation. Although qualitative traits were not included in the dimensionality reduction analysis, we also investigated the association of the different qualitative traits with the main axes of trait variation. Prior to the analyses, we excluded outliers and standardized the data. Due to the high sensitivity of dimensionality reduction to outliers, we excluded values within the 2.5th--97.5th percentile range (\protect\hyperlink{ref-legendre2012}{Legendre \& Legendre, 2012}), and thus our final dataset had 1,236 species. Then, we log transformed the variables to reduce the influence of outliers and z-transformed (X= 0, SD=1) so that all variables were within the same numerical range. We performed the pPCA using the function \emph{phyl.pca} from the package \emph{phytools} (\protect\hyperlink{ref-revell2012}{Revell, 2012}) with the method lambda (\(\lambda\)) that calculates the phylogenetic correlation between 0 (phylogenetic independence) and 1 (shared evolutionary history) and we implemented the mode covariance because values for each variables were on the same scale following transformation (\protect\hyperlink{ref-abdi2010}{Abdi \& Williams, 2010}). Moreover, to corroborate that our imputation of missing values did not affect our results, we conducted a pPCA on the full dataset without missing values (Fig. S3). We found little difference between the explained variance with the imputed dataset (51.08\%) and the dataset without missing values (52.87\%). In addition, the loadings on each principal component had a similar contribution and correlation patterns, with the exception of plant height which showed slight variations between the imputed and non-imputed dataset. Finally, we conducted an additional phylogenetic informed principal component analysis for the subset of species with pollen and nectar quantity. For this, we included all quantitative traits considered in the main pPCA plus pollen grains and microlitres of nectar per flower.

@@

We calculated the phylogenetic signal of the different quantitative traits on the imputed dataset with the full set of species (N = 1,506) with the package \emph{phytools} (\protect\hyperlink{ref-revell2012}{Revell, 2012}) and we used Pagel's \(\lambda\) as a measurement of the phylogenetic signal. However, for pollen and nectar traits, phylogenetic signal was calculated only on the subset of species that had quantitative information for these traits (N = 755).

@@

Analyses were conducted on the subset of 60 weighted networks sampled in a unique flowering season and site, which included 556 plant and 1,126 pollinator species. These networks were analysed in their qualitative and quantitative form. First, we analysed the binary version of these weighted networks with presence-absence information that assumes equal weight across interactions. Second, we analysed the untransformed weighted networks with the number of visits of floral visitors to individual flowers that account for the intensity of the interaction. Although floral visitors are not always pollinators and number of visits does not consider each pollinator species efficiency (\protect\hyperlink{ref-ballantyne2015}{Ballantyne et al., 2015}), the number of visits can provide valuable information of the contribution of floral visitors to pollination (\protect\hyperlink{ref-vazquez2005}{Vázquez et al., 2005}, \protect\hyperlink{ref-vazquez2012}{2012}). In total, our network dataset (excluding meta-webs and non-weighted networks) included 2,256 interactions of bees with plants, 1,768 non-syrphid-Diptera interactions, 845 syrphids interactions, 437 Lepidoptera interactions, 432 Coleoptera interactions and 362 non-bee-Hymenoptera interactions. Sampling methods varied across networks but this was accounted for in analyses by considering them in the random effects of the modelling process. All analyses were conducted in R version \emph{4.0.3}.

@@

We used Bayesian modelling (see below for details) to explore the effect of floral visitor groups and the main axes of trait variation (pPCA with imputed dataset) on both qualitative and quantitative floral interactions per plant species. For this, we divided floral visitors into six main guilds that differ in life form, behaviour and are likely to play a similar ecological role: (i) bees (Hymenoptera-Anthophila), (ii) non-bee-Hymenoptera (Hymenoptera-non-Anthophila), (iii) syrphids (Diptera-Syrphidae), (iv) non-syrphid-Diptera (Diptera-non-Syrphidae), (v) Lepidoptera and (vi) Coleoptera. Moreover, because the guild of bees was the most represented group with 2,256 records and had the highest frequency of visits, we also explored presence-absence of the interaction and number of visits of the main bee families (Andrenidae, Apidae, Colletidae, Halictidae and Megachilidae) on the trait space. In addition, we found that \emph{Apis mellifera} was the floral visitor with the largest proportion of records counted (7.55\% of the total). This finding is consistent with previous research showing that \emph{A. mellifera} was the most frequent floral visitor in a similar dataset of 80 plant-pollinator networks in natural ecosystems (\protect\hyperlink{ref-hung2018}{Hung et al., 2018}). Hence, to control for the effect of \emph{A. mellifera} on the observed visitation patterns of bees, we conducted an analogous analysis with presence-absence of the interaction and number of visits excluding \emph{A. mellifera}. We found that \emph{A. mellifera}, was partly driving some of the observed trends on PC1 (Fig. S4). However, we did not detect major differences on PC2 and PC3.

We implemented Bayesian generalized linear mixed models using the R package \emph{brms} (\protect\hyperlink{ref-burkner2017}{Bürkner, 2017}). We modelled presence-absence of observed interactions and number of visits as a function of the main axes of plant trait variation and their interactions with floral visitor guilds (e.g., number of visits \textasciitilde{} PC1 x FGs + PC2 x FGs + PC3 x FGs). Because we were interested in possible differences in the visitation patterns among floral visitors groups to plants with different strategies, we included interactions between the main axes of trait variation (PC1, PC2 and PC3) and the different floral visitor guilds. We added a nested random effect of networks nested within the study system to capture the variation in networks among studies and within networks. Moreover, we included the phylogenetic covariance matrix as a random factor due to the possible shared evolutionary histories of species and therefore lack of independence across them. We specified for presence-absence of interaction and number of visits a Bernoulli and a zero inflated negative binomial distribution, respectively. The models were run with non or very weakly informative informative priors from the brm function so they have neglible influence on the results (\protect\hyperlink{ref-burkner2017}{Bürkner, 2017}), 3,000 iterations and with previous 1,000 warm up iterations. We set delta (\(\Delta\)) to 0.99 to avoid divergent transitions and visualized the posterior predictive checks with the function \emph{pp\_check} using the \emph{bayesplot} package (\protect\hyperlink{ref-gabry2019}{Gabry et al., 2019}).
